---
title: "Lab 10: Moment-Generating Functions, Method of MGFs, and Order Statistics"
subtitle: 'MTH 561: Mathematical Statistics I'
date: "Last updated: `r Sys.Date()`"
output: 
  html_document:
      toc: TRUE
      toc_depth: 2
      toc_float: TRUE
      css: style_MTH561.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)

# library(htmltools)
# 
# footer <- div(
#   style = "font-size: 8pt; text-align: right; width: 100%; padding-right: 12pt;", 
#   span(class = "pageNumber"), "/", span(class = "totalPages")
# )
# 
# pagedown::chrome_print(
#   "MTH561_Lab01.html", 
#   options = list(
#     displayHeaderFooter = TRUE, 
#     footerTemplate = format(footer, indent = FALSE),
#     marginTop = 0.4,
#     marginBottom = 0.4
#   )
# )
```

# Sections 3.9 and 4.9: Moment-Generating Functions

1. (Q3.147) If $Y$ has a geometric distribution with probability of success $p$, show that the moment-generating function for $Y$ is 

$$m(t) =\frac{pe^t}{1-qe^t}$$

where $q=1-p$.

2. (Q3.149-3.152, 4.140). Use the uniqueness of moment-generating functions to give the distribution of a random variable with the following moment-generating functions.

    a) $$m(t) =(0.6e^t+0.4)^3$$
    b) $$m(t) = \frac{0.3e^t}{1-0.7e^t}$$
    c) $$m(t) = (0.7e^t + 0.3)^{10}$$
    d) $$m(t)= e^{6(e^t-1)}$$
    e) $$m(t) =(1-4t)^{-2}$$
    f) $$m(t) = 1/(1-3.2t)$$
    g) $$m(t) = e^{-5t+6t^2}$$

3. (Q3.155) Let $m(t) = (1/6)e^t + (2/6)e^{2t} + (3/6)e^{3t}$. Find:

    a) $E(Y)$
    b) $V(Y)$
    c) The distribution of $Y$

4. (Q3.156) Suppose that $Y$ is a random variable with moment-generating function $m(t)$.

    a) What is $m(0)$?
    b) If $W=3Y$, show that the moment-generating function of $W$ is $m(3t)$.
    c) If $X=Y-2$, show that the moment-generating function of $X$ is $e^{-2t}m(t)$.
    
5. (Q4.141) If $\theta_1 < \theta_2$, derive the moment-generating function of a random variable that has a uniform
distribution on the interval $(\theta_1,\theta_2)$.

6. (Q4.145) A random variable $Y$ has the density function below.

    a) Find $E(e^{3Y/2})$.
    b) Find the moment-generating function for $Y$.
    
$$f(y)=\begin{cases}
e^{y} & y<0\\
0 & else
\end{cases}$$
    
# Section 6.5: Method of Moment-Generating Functions

1. (Q6.39) Consider two electronic components that operate independently, each with a length of life following an exponential distribution with mean 1. Use the method of moment-generating functions to find the density function for the average life of the two components.

2. (Q6.40) Suppose that $Z_1$ and $Z_2$ are independent, standard norrmal random variables. Find the density function of $U=Z_1^2 + Z_2^2$.

3. (Q6.41) Let $Y_1, ..., Y_n$ be independent normal random variables, each with mean $\mu$ and variance $\sigma^2$. Let $a_1, ..., a_n$ denote known constants. 

    a) Find the density function of the linear combination $U = \sum_{i} a_i Y_i$.
    b) Find the density function of the sample mean, $\bar{Y} = \frac{1}{n}\sum_{i} a_i Y_i$.
    
4. (Q6.49) Let $Y_1$ be a binomial random variable with $n_1$ trials and probability of success given by $p$. Let $Y_2$ be another binomial random variable with $n_2$ trials and probability of success also given by $p$. If $Y_1$ and $Y_2$ are independent, find the probability function of $Y_1 + Y_2$. 

# Section 6.7: Order Statistics

1. (Q6.73) Let $Y_1$ and $Y_2$ be independent and uniformly distributed over the interval (0,1). Find:

    a) The probability density function of $U = max(Y_1, Y_2)$.
    b) $E(U)$ and $V(U)$ _without integrating_.
    
2. Let $Y_1, Y_2, ..., Y_n$ be independent and uniformly distributed over the interval (0,1). Find:

    a) The probability density function of $U = max(Y_1, Y_2, ..., Y_n)$.
    b) $E(U)$ and $V(U)$ _without integrating_.
    
3. (Q6.81) Let $Y_1, ..., Y_n$ be independent random variables, each distributed exponentially with mean $\beta$.

    a) Show that $Y_{(1)} =min(Y_1, Y_2, ..., Y_n)$ has an exponential distribution. Find its mean parameter.
    b) Argue that $Y_{(n)} =max(Y_1, Y_2, ..., Y_n)$ does _not_ an exponential distribution. Find its distribution - is it one that we know?
    
```{r, echo=TRUE}
# You may use the results of this simulation to support your answer

N <- 1000
n <- 100
lambda <- 1

# Create an "empty vector" to store our results
max <- vector(length=N)

for(i in 1:N){
  # Generate a random sample with n=100 observations following an Exponential(lambda) distribution
  sample <- rexp(n=n, rate=lambda)
  # Calculate the max of each column, and store it in max vector
  max[i] <- max(sample)
}

# Make a histogram of the sample maximums
hist(max)

```
	
---

# Problems to submit

__Question 1__: If $Y$ is a random variable with moment-generating function $m(t)$ and if $W$ is given by $W=aY+b$, show that the moment-generating function of $W$ is $e^{tb} m(at)$. (This result applies to both discrete and continuous $Y$, but make your argument for the continuous case.)

__Question 2__: Suppose that $Y$ has a gamma distribution with $\alpha=n/2$ for some positive integer $n$ and $\beta$ equal to some specified value. Show that $W=2Y/\beta$ has a $\chi^2$ distribution with $n$ degrees of freedom.

__Question 3__: Let $Y_1, ...Y_n$ be independent random variables, each distributed $Beta(\alpha=\beta=2)$. Find:

a) The probability distribution function of $Y_{(n)} = max(Y_1, ..., Y_n)$.
b) The density function of $Y_{(n)}$.
c) A general expression for $E(Y_{(n)})$ _without integrating_.